---
title: "data.table"
author: "Aaron Schroeder"
date: ""
output: 
  html_document: 
    keep_md: yes
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(data.table)

getwd()

if (interactive()) {
  data_file <- "data/GB_full.csv.zip"
} else {
  data_file <- "../../data/GB_full.csv.zip"
}
data_file
```
### Load Data
```{r cache=TRUE}
!file.exists(data_file)

if (!file.exists(data_file)) {
    download.file("http://download.geonames.org/export/zip/GB_full.csv.zip", data_path)
}
DT <- fread(unzip(data_path))
```

```{r}
head(DT)
tail(DT)
```

### How to subset columns (note use of .(column list) syntax)
```{r}
#subsetting columns
sub_columns <- DT[, .(V2,V3,V4)]
head(sub_columns)
```

### How to subset rows (note no need for following comma)
```{r}
#subsetting rows
sub_rows <- DT[V4 == "England" & V3 == "Beswick"]
head(sub_rows)
```

### How to order variables in ascending or descending order (note use of "-" to connote descending order)
```{r}
#ordering columns
dt_order <- DT[order(V4, -V8)]
head(dt_order)
```

### How to add a column
```{r}
#add a new column - here we are adding columns V10 and V11 to create new column V_New
DT[, V_New := V10 + V11]
```

### How to update a columns value based on it's current value
```{r}
#update row values - here we change the value to "Abr City" where it equals "Aberdeen City"
DT[V8 == "Aberdeen City", V8 := "Abr City"]
```

### How to update a columns value based on another column's value
```{r}
#update row values - here we change the value of column V8 to "Aberdeen City" where column V6 equals "Aberdeen City"
DT[V6 == "Aberdeen City", V8 := "Aberdeen City"]
```

### How to delete a column
```{r}
#delete a column - here we delete 2 columns
DT [,c("V6","V7") := NULL ]
```

### You can Chain Commands Together!
```{r}
DT[V8 == "Aberdeen City", V8 := "Abr City"][, V_New := V10 + V11][,c("V6","V7") := NULL]
```

### Use functions (e.g. aggregates) on a column
```{r}
#compute the average - note you must remove NAs for an aggreagte function to work
DT[, .(average = mean(V10, na.rm = TRUE))]
```

### Use functions (e.g. aggregates) on a column with grouping
```{r}
#compute the average
DT[, .(average = mean(V10, na.rm = TRUE)), by = V4]
```

### Get counts by grouping
```{r}
#compute the count
DT[, .N, by = V4]
```
