---
title: "README"
author: "Daniel Chen"
date: ""
output:
  md_document:
    variant: markdown_github 
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

To prepare the data example I will use the `mtcars` dataframe,
and create a new column, `cars` by passing in the `row.name`

```{r}
mtcars$cars <- row.names(mtcars)
```


# Functions

```{r}
my_mean <- function(){
  print(3)
}

my_mean()
```

```{r}
my_mean <- function(x, y){
  print(3)
}

my_mean()

my_mean(10, 20)
```


## Functions will automaticall return the last caculated value
```{r}
my_mean <- function(x, y) {
    (x + y) / 2
}
```

## Explicitly return a value in a function
```{r}
my_mean <- function(x, y) {
    return((x + y) / 2)
}
```

```{r}
my_mean <- function(x, y) {
    val <- (x + y) / 2
    return(val)
}
```

# Functions on data

The summary function worked on our data column-by-column.

You can write your own function that work column-by-column, or row0-by-row by using `apply`


## Visually check your function

When starting to write a function that you will apply on your data,
it's good to write the function on its own with the parameters you will use to visually inspect it

```{r}
# create a function that calculates the ratio of a cyl_val and wt_val
# append this ratio to a car_name separated by a dash
rename_car <- function(cyl_val, wt_value, car_name) {
    cyl_wt_ratio <- as.numeric(cyl_val) / as.numeric(wt_value)
    new_name <- paste(car_name, '-', cyl_wt_ratio)
    return(new_name)
}

rename_car(10, 20, 'pizaaaaaaaaa')
```

## Preparing your function for an `apply`

The way apply works is the first parameter is the dataframe you will work on

The second parameter, `MARGIN` will tell apply to work row-by-row (`1`)
or column-by-column (`2`).

The last parameter is the function you want to apply,
note that you do not need the round brackets when you pass in functions into apply,
just the function name.

To convert your function to be `apply` ready, you replace the parameters
of your function with a new parameter, e.g., `row_data` or `col_data`.
then within the function, you will use the subsetting methods to subset the rows or columns your function needs

```{r}
rename_car <- function(row_data) {
    cyl_val <- row_data['cyl']
    wt_value <- row_data['wt']
    car_name <- row_data['cars']

    cyl_wt_ratio <- as.numeric(cyl_val) / as.numeric(wt_value)

    new_name <- paste(car_name, '-', cyl_wt_ratio)
    return(new_name)
}
```

with your new function, you can then apply it to your data

```{r}
apply(mtcars, 1, rename_car)
```

You can see this returns a vector of values, but we want to assign this to a new column

```{r}
mtcars$car_ratio <- apply(mtcars, 1, rename_car)

head(mtcars)
```


# Common mistakes

What went wrong when I was live coding?

Here's the code I originally typed

```{r}
rename_car <- function(cyl_val, wt_value, car_name) {
    cyl_wt_ratio <- as.numeric(cyl_val) / as.numeric(wt_value)
    new_name <- paste(car_name, '-', cyl_wt_ratio)
    return(new_name)
}

```

The first time I `apply`ified the function I wrote this:

```{r}
rename_car <- function(row_data, cyl_val, wt_value, car_name) {
    cyl_wt_ratio <- as.numeric(cyl_val) / as.numeric(wt_value)
    new_name <- paste(car_name, '-', cyl_wt_ratio)
    return(new_name)
}

```

Because the way apply works is that the row (or column) data all gets passed into the
first argument of the function, but the output was 'strange'/wrong.

I attempted to fix this by subsetting the 'columns' that the row is passed in.

```{r, error=TRUE}
rename_car <- function(row_data) {
    cyl_val <- row_data[, 'cyl']
    wt_value <- row_data[, 'wt']
    car_name <- row_data[, 'cars']

    cyl_wt_ratio <- cyl_val / wt_value

    new_name <- paste(car_name, '-', cyl_wt_ratio)
    return(new_name)
}

apply(mtcars, 1, rename_car)
```

That didn't work because the row of data got passed in.
This row or `vector` can't be subset by using the `[, ]` notation (not the comma)
because there is no way you can subset both a row and column on a 1-dimentional thing (which a row/vector is)

So I had to remove the comma from the subsetting

```{r, error=TRUE}
rename_car <- function(row_data) {
    cyl_val <- row_data['cyl']
    wt_value <- row_data['wt']
    car_name <- row_data['cars']

    cyl_wt_ratio <- cyl_val / wt_value

    new_name <- paste(car_name, '-', cyl_wt_ratio)
    return(new_name)
}

apply(mtcars, 1, rename_car)
```

Next problem was this `non-numeric argument to binary operator` error.

It tells me that the error occurs when trying to calculate `cyl_val/wt_value`.

the `binary operator` the error message is refering to is the division `/`,
and it is telling me that values passed into the division is not a `numeric` value.

We fix this by forcing the values to be `numeric` by using the `as.numeric` function

```{r, error=TRUE}
rename_car <- function(row_data) {
    cyl_val <- row_data['cyl']
    wt_value <- row_data['wt']
    car_name <- row_data['cars']

    cyl_wt_ratio <- as.numeric(cyl_val) / as.numeric(wt_value)

    new_name <- paste(car_name, '-', cyl_wt_ratio)
    return(new_name)
}
```

you can also do the conversion at the very beginning

```{r, error=TRUE}
rename_car <- function(row_data) {
    cyl_val <- as.numeric(row_data['cyl'])
    wt_value <- as.numeric(row_data['wt'])
    car_name <- row_data['cars']

    cyl_wt_ratio <- cyl_val / wt_value

    new_name <- paste(car_name, '-', cyl_wt_ratio)
    return(new_name)
}
```

In R (and statistical datasets), we know each column will be the same data type (character, numeric, etc)

But in statistical datasets, rows can be of mixed types.

What R does in the apply is that when the row data is passed,
it converts everything into the data type that is common to all the row values, a character vector (aka string)

which is why we need to manually convert the `cyl` and `wt` values into numerics.
