---
title: "r_sql"
author: "Daniel Chen"
date: ''
output:
  html_document: 
    keep_md: yes
    toc: yes
  html_notebook: default
---

```{r setup, include=FALSE}

SUPER_SECRET_PASSWORD <- ""


knitr::opts_chunk$set(echo = TRUE)
```

# Connecting to a SQL database

```{r, eval=FALSE}
library(sdalr)
library(DBI)
library(RPostgreSQL)

con <- con_db(dbname = 'arlington', user = 'chend')
```

```{r, echo=FALSE}
# DELETE PASSWORD WHEN COMMITING
library(sdalr)
library(DBI)
library(RPostgreSQL)
con <-  con_db(dbname = 'arlington', user = 'chend', pass = SUPER_SECRET_PASSWORD)
```

```{r}
con
```

# Getting (Querrying) data from database in R

```{r}
df <- dbGetQuery(con, "")
head(df)
```

You can also use an `sql` chunk
```
```{sql connection=con}
SELECT * FROM fire.medic_unit_movement_summary_2013;
```
```

```{sql connection=con}
SELECT * FROM fire.medic_unit_movement_summary_2013;
```

# Creating Tables

Everyone should have a database named their pid.

```{r, eval=FALSE}
con <- con_db(dbname = 'chend', user = 'chend')
```

```{r, echo=FALSE}
con <-  con_db(dbname = 'chend', user = 'chend', pass = SUPER_SECRET_PASSWORD)
```

```{r, results='hide'}
# delete a table (you have permissino to do)
dbSendQuery(con, "DROP TABLE customers;")

# create a new table
dbSendQuery(con, "CREATE TABLE customers (customer_no INTEGER, first_name TEXT, last_name TEXT);")

# add row information into a table
dbSendQuery(con, "INSERT INTO customers (customer_no, first_name, last_name) VALUES (1, 'MC', 'Hammer')")
dbSendQuery(con, "INSERT INTO customers (customer_no, first_name, last_name) VALUES (2, 'Hammer', 'Time')")
```


```{r}
dbGetQuery(con, "SELECT * from customers;")
```

```{r}
# save results to an R data frame

df <- dbGetQuery(con, "SELECT * from customers;")
df
```

# Get data in and out of chunks

When working with chunks that use different language engines, you will need a way to pass
R values into the engine, and get values back into the R environment

## Get R data into chucnks

To get an `R` variable into a chunk, you prepend the R variable by a `?`, e.g., `?x` if you want to use the variable `x` in the chunk

```{r}
x <- 'Time'
```

```
```{sql connection=con}
SELECT * FROM ?x;
```
```

```{sql connection=con}
SELECT * FROM customers WHERE last_name = ?x;
```

## Get chucnk values into R

You can get values and results back into the R enviornment by using the `output.var` chunk option

```
```{sql connection=con, output.var=df_sql}
SELECT * FROM ?x;
```
```

```{sql connection=con, output.var='df_sql'}
SELECT * FROM customers WHERE last_name = 'Hammer';
```

In R:

```{r}
df_sql
```

